Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
add edits to remove horrible dnf lenses

\subsection{Symmetric DNF Lenses}
Symmetric DNF lenses are an intermediate synthesis target for \GreedySynth.
Because of the lack of a type equivalence rule or a composition lens, the set of
possible lenses with type $\overline{\DNFRegex} \Leftrightarrow
\overline{\DNFRegexAlt}$ is quite small. Consequently, an algorithm for
synthesizing such lenses can be driven primarily by the type of the lenses.

\saz{The following two paragraphs are pretty similar (was there a bad merge at
  some point?) and very low level.  I also
  could not really follow either of the examples at all.  If we
  are looking for things to cut, I would propose dropping most of the content of
  the next two paragraphs, merging the remainder, and skipping pretty directly to the paragraph
  that starts with ``While DNF lenses...''}
As an example, consider the symmetric DNF lens:
$(\SDNFLensOf{(1,1,\SSQLens_1)\DNFLSep (1,2,\SSQLens_2)}
,\ListOf{2},\ListOf{1;2})$, synchronizing data between $\DNFOf{\Sequence_1}$ and
$\DNFOf{\SequenceAlt_1 \DNFSep \SequenceAlt_2}$. This lens merges both
$\SequenceAlt_1$ and $\SequenceAlt_2$ into $\Sequence_1$, as they both are
involved in a lens with $\Sequence_1$. When performing a \CreateR, $\SSQLens_2$
is used. When performing a \CreateL on a string in $\Sequence_1$, $\SSQLens_1$
is used, and when performing a \CreateL on a string in $\Sequence_2$,
$\SSQLens_2$ is used. More generally, a symmetric DNF Lens \SDNFLens{} consists
of three components. First is the mapping component. The mapping component is a
list of triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component of the triple, \SSQLens{}, is a sequence lens that maps between
the two sequences specified by $i$ and $j$. The second component is a list of
integers $c$. The integer, $c_i$ at position $i$ states that when performing a
$\CreateR$ on a string matching sequence $\Sequence_i$, it will use the sequence
lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
component is a list of integers $d$ that follows the same pattern as the second
component, but maps from right to left.

As an example, consider the symmetric sequence lens:
$(\SSQLensOf{(1,2,\SAtomLens_1)} ,\ListOf{}
,\ListOf{(\StringAlt_1',\AtomAlt_1)})$, synchronizing data between
$\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \String_1}$ and
$\SequenceOf{\StringAlt_0 \SeqSep \AtomAlt_1 \SeqSep \StringAlt_1 \SeqSep
  \AtomAlt_2 \SeqSep \StringAlt_2}$. This lens maps $\Sequence_1$ to
$\SequenceAlt_2$, and leaves $\AtomAlt_1$ disconnected. On a \CreateR, the
string created is a concatenation of $\StringAlt_1'$ with the default from
$\SAtomLens_1$. More generally, a sequence lens \SSQLens{} consists of three
components. First is the mapping component. The mapping component is a list of
triples $(i,j,\SAtomLens)$. In this, $i$ and $j$ refer to atom $i$ and atom $j$
on the left and right sequences, respectively. The third component of the
triple, \SAtomLens{}, is an atom lens that maps between the two atoms specified
by $i$ and $j$. The second component is a list of strings, $\String$. The
string, $s_i$ at position $i$ states that when performing a $\CreateL$ on a
string matching atom $\Atom_i$, it will do one of two things. If $\Atom_i$ is
mapped to in the mapping component, then performing the atom's $\CreateL$ will
actually $\PutL$ into that string. If $\Atom_i$ is not mapped to, then
$\String_i$ will be created as a representative of $\Atom_i$.

While DNF lenses are useful because they permit an efficient greedy synthesis
procedure, the particulars of their typing and semantics are not the primary
intellectual contributions of our work. The details of the typing and semantics
are included in the supplementary material present in the appendix.

\paragraph*{Information-Theoretic Cost Metric}
Much like we defined a syntactic means to find the expected number of bits to
recover string from a synchronized string in another format for simple symmetric
string lenses, we do the same for symmetric DNF lenses. Unlike simple symmetric
string lenses, DNF lenses do not have composition, so the entropy can be defined
over all terms. $\REntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
expected number of bits required to recover a string in $\DNFRegex$ from a
synchronized string in $\DNFRegexAlt$.

\[
  \begin{array}{rl}
    & \REntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
    & \hspace*{1.21em}\Given
      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
      ,\ListOf{c_1;\ldots;c_q}
      ,\ListOf{d_1;\ldots;d_r})\\
    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
    & \text{where } H_j = (0,\Sigma_{\SetOf{k\SuchThat j_k = j}}
      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j})
  \end{array}
\]

@@

In this formulation, data unique to each side are stored in the complement. When
one format is edited, the \PutR or \PutL function stitches together the edited
data with data stored in the complement. The $init$ element is the initial value
of $C$ and specifies default behavior when data is missing. For instance, to
implement the scenario in Figure~\ref{fig:minimized-representations}, the
complement would consist of a list of pairs of salary and company name.
Classical symmetric lenses satisfy the following equational laws.
\begin{equation}
  \tag{\PutRL}
  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'''')} = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c'')} = (y,c')}
\end{equation}

@@

\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.

@@

\[
  \inferrule*
  {
    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
    \ldots\\
    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
    \String_1 \in \Atom_1\\
    \ldots\\
    \String_q \in \Atom_q\\\\
    \StringAlt_1 \in \AtomAlt_1\\
    \ldots\\
    \StringAlt_r \in \AtomAlt_r\\\\
    i_x = i_y \BooleanImplies x = y\\
    j_x = j_y \BooleanImplies x = y
  }
  {
    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
      \ldots\SeqLSep
      (i_p,j_p,\SAtomLens_p)}
    ,\ListOf{\String_1;\ldots;\String_q}
    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
    \OfType
    \SequenceOf{\String_0'''' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
      \SeqSep \String_q'}
    \Leftrightarrow
    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
  }
\]

>>>>

\subsection{Symmetric DNF Lenses}
Symmetric DNF lenses are an intermediate synthesis target for \GreedySynth.
There are many fewer symmetric DNF lenses than symmetric regular lenses. In
fact, if one does not use the star-unrolling axioms, there are only finitely
many DNF lenses of a given type (though there are still many more symmetric DNF
lenses than there are DNF bijective lenses).

Intuitively, a symmetric DNF lens $sdl$ is a union of symmetric sequence lenses,
$ssql_1 \ldots ssql_p$, a symmetric DNF sequence lens is a concatenation of
symmetric atom lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is
an iteration of a symmetric DNF lens.

Just as we analyzed the information content of ordinary regular expressions, we
can analyze the information content of DNF regular expressions. As before, we
use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
of the expected amount of information required to recover a string in
$\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
calculate bounds of the expected amount of information required to recover a
string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
$\SDNFLens$.

The details of these definitions are syntactically tedious, but not
intellectually difficult. We elide them here but include them in the appendix.

@@

In this formulation, data unique to each side are stored in the complement. When
one format is edited, the \PutR or \PutL function stitches together the edited
data with data stored in the complement. The $init$ element is the initial value
of $C$ and specifies default behavior when data is missing. For instance, to
implement the scenario in Figure~\ref{fig:minimized-representations}, the
complement would consist of a list of pairs of salary and company name.
Classical symmetric lenses satisfy the following equational laws.
\begin{centermath}
\begin{array}[b]{l@{\qquad \; \qquad}l}
  \begin{mathprooftree}
    \AxiomC{$\PutRSymOf{(x,c)} = (y,c')$}
    \UnaryInfC{$\PutLSymOf{(y,c')} = (x,c')$}
  \end{mathprooftree}
&
  \begin{mathprooftree}
    \AxiomC{$\PutLSymOf{(y,c)} = (x,c')$}
    \UnaryInfC{$\PutRSymOf{(x,c')} = (y,c')$}
  \end{mathprooftree}
\end{array}
\end{centermath}

@@

\paragraph*{Symmetric DNF Lenses: Syntax, Cost, Typing, and Semantics}

\paragraph*{Information-Theoretic Cost Metric}
Much like we defined a syntactic means to find the expected number of bits to
recover string from a synchronized string in another format for simple symmetric
string lenses, we do the same for symmetric DNF lenses. Unlike simple symmetric
string lenses, DNF lenses do not have composition, so the entropy can be defined
over all terms. $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
expected number of bits required to recover a string in $\DNFRegex$ from a
synchronized string in $\DNFRegexAlt$.

\[
  \begin{array}{rl}
    & \LEntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
    & \hspace*{1.21em}\Given
      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
      ,\ListOf{c_1;\ldots;c_q}
      ,\ListOf{d_1;\ldots;d_r})\\
    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
    & \text{where } H_j = (0,\Sigma_{\SetOf{k\SuchThat j_k = j}}
      \LEntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j})
  \end{array}
\]

The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.

@@

\[
  \inferrule* {
    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
    \ldots\\
    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
    \String_{c_1} \in \Atom_{c_1}\\
    \ldots\\
    \String_{c_{q'}} \in \Atom_{c_{q'}}\\\\
    \StringAlt_{d_1} \in \AtomAlt_{d_1}\\
    \ldots\\
    \StringAlt_{d_{r'}} \in \AtomAlt_{d_{r'}}\\\\
    i_x = i_y \BooleanImplies x = y\\
    j_x = j_y \BooleanImplies x = y } {
    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep \ldots\SeqLSep
      (i_p,j_p,\SAtomLens_p)}
    ,\ListOf{(\String_{c_1},\Atom_{c_1});\ldots;(\String_{c_{q'}},\Atom_{c_{q'}})}
    ,\ListOf{(\StringAlt_{d_1},\Atom_{d_1});\ldots;(\StringAlt_{d_r},\Atom_{d_r})})\\
    \OfType \SequenceOf{\String_0'''' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep
      \Atom_{q'} \SeqSep \String_{q}'} \Leftrightarrow \SequenceOf{\StringAlt_0'
      \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep
      \StringAlt_{r}'} }
\]
